---
title: "Lab1"
author: "HuanChen"
date: "2019/3/31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(survival)
library(lmtest)
```

survival: Contains the core survival analysis routines, including definition of Surv objects, Kaplan-Meier and Aalen-Johansen (multi-state) curves, Cox models, and parametric accelerated failure time models.

lmtest: A collection of tests, data sets, and examples for diagnostic checking in linear regression models. Furthermore, some generic tools for inference in parametric models are provided.


# 1. Overdispersion
```{r}
## Read Crabs dataset into R 
Crabs <- read.table("http://www.stat.ufl.edu/~aa/glm/data/Crabs.dat",header = T)
## Attach the Crabs data
attach(Crabs)
## Take a look at Crabs
head(Crabs)
dim(Crabs)
```

A data frame with 173 observations on the following 5 variables.

* color
    + A factor with levels 1 = light medium, 2 = medium, 3 = dark medium, 4 = dark.
* spine
    + A factor with levels 1 = both good, 2 = one worn or broken, 3 = both worn or broken.
* width
    + Crab carapace width in cm.
* satell
    + Number of satellites.
* weight
    + Crab weight in in kg.

```{r}

## Fit Poisson Log linear Regression model without considering overdispersion
fit.pois <- glm(y ~ weight, family=poisson) # ML Poisson loglinear
summary(fit.pois)

## Pearson Residual
Pr <- sum(residuals(fit.pois, type="pearson")^2) / fit.pois$df.residual
Pr  # estimate for overdispersion parameter
## Compute Residual Deviance
Rr <- fit.pois$deviance / fit.pois$df.residual
Rr

## Fit Poisson Log-Linear Regression with quasi likelihood
fit.quasi <- glm(y ~ weight, family=quasi(link="log",variance="mu"))
summary(fit.quasi)
```

# 2. Model Diagnostics
```{r}
## Leverage
hatvalues(fit.quasi) %>% range()
plot(1:nrow(Crabs),hatvalues(fit.quasi),pch=19)

## Outlier Using Residuals
plot(fit.quasi, 3)  # standardized deviance residuals

## Cook's Distance
cooks.distance(fit.quasi) %>% range
plot(fit.quasi, 4)  # Cook's distance
plot(fit.quasi, 5)  # std. Pearson residual vs leverage

```

# 3. Conditional Logistic regression
```{r}
## The Logan dataset
head(logan, 20)
dim(logan)
```

* Intergenerational occupational mobility data with covariates
* occupation
    + subject's occupation, a factor with levels farm, operatives, craftsmen, sales, and professional
* focc
    + father's occupation
* education
    + total years of schooling, 0 to 20
* race
    + levels of non-black and black

```{r}
resp <- levels(logan$occupation)
n <- nrow(logan)
indx <- rep(1:n, length(resp))
logan2 <- data.frame(logan[indx,],
                     id = indx,
                     tocc = factor(rep(resp, each=n)))
logan2$case <- (logan2$occupation == logan2$tocc)
logan2 <- logan2[order(logan2$id),]

## Show dataset for first three strata
logan2[logan2$id %in% c(1,2,3), ]

## Analysis clogit
res.clogit <- clogit(case ~ tocc + strata(id), logan2)
summ.clogit <- summary(res.clogit)
summ.clogit

## drop in deviance test
set.seed(20180327)
logan2 <- cbind(logan2, new = rnorm(4190))
logan2 %>% head  # random noise added

res.clogit.1 <- clogit(case ~ tocc + new + strata(id), logan2)
summ.clogit.1 <- summary(res.clogit.1)
summ.clogit.1

lrtest(res.clogit, res.clogit.1)
```


